---
title: "STT 3851 Project 2 Final"
author: "Kassidy Borum, Adam Seagle, Jordan Greene"
date: "5/06/2020"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 1
    number_sections: false
bibliography: [packages.bib]    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warning = FALSE)
library(readxl)
library(tidyverse)
library(leaps)
library(glmnet)
library(pls)
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'tidyverse', 'base', 'readxl',
  'corrplot'
), 'packages.bib')
```

```{r data, include=FALSE}
house <- read_xlsx("Housing.xlsx")
```


#A: First Candidate Model 
```{r}
#Final Model from Project 1
model1 <- lm(price ~ bedrooms + bath + garagesize + status + 
                       elem + size + lot + bedrooms:bath, house)
summary(model1)
```

# B: Second Candidate Model
```{r}
names(house)
dim(house)
sum(is.na(house))
```


```{r}
reg.fit.full <- regsubsets(price ~ ., data=house, method="forward", really.big = TRUE)
reg.summary <- summary(reg.fit.full)
reg.summary
names(reg.summary)
```

  - Based on the summary of the full regfit model, the best predictors are `lot`, `garagesize`, `elemedison`, and `elemharris`
  
```{r}
reg.summary$rsq
```
  - You can see the $R^2$ value increases from 12.84% with one variable to 53.61% when we use all the varibles. 
  
  

```{r}
reg.summary$bic

par(mfrow=c(2,2))
plot(reg.summary$rsq, xlab = "Number of Variables", ylab = "R-squared",type="l")
which.max(reg.summary$rsq)

plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adjusted R-squared",type="l")
which.max(reg.summary$adjr2)

points(9, reg.summary$adjr2[9], col = "green", cex = 2, pch = 20)

plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)

points(6, reg.summary$cp[6], col = "purple", cex = 2, pch = 20)
which.min(reg.summary$bic)

plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")

points(6, reg.summary$bic[6], col = "pink", cex = 2, pch = 20)
```

  - The Adjust $R^2$ is shows that 9 predictors is the best. 
  
```{r}
#view(house)
plot(reg.fit.full,scale="r2")
plot(reg.fit.full,scale="adjr2")
plot(reg.fit.full,scale="Cp")
plot(reg.fit.full,scale="bic")
coef(reg.fit.full, 9)
```
  - The significant coeffcients are `size`, `lot`, `bedrooms`, `status`, and `elementary`. 
 
## Model Summary {-}

```{r}
model2 <- lm(price ~ size + lot + bedrooms + garagesize + status + elem, data = house)
summary(model2)
```
  - This model has a slightly smaller adjust R-squared value (0.457) in comparison to our intial model from project 1 (.4797)
    - `status` and `elem` have different cateogrical values:
        - `status`: statuspen, statussld, 
        - `elem`: elemcrest, elemedge, elemedison, elemharris, elemparker
    - only statussld and elemedison are seen to be statically significant, while elemharris is slightly above the desire p-value of 0.05. 
    
#C: Train/Test Split

```{r}
set.seed(1)
index <- sample(nrow(house), nrow(house) * 0.50, replace = FALSE)

train <- house[index,]
test <- house[-index,]
```

#D: Third Candidate Model

```{r}
reg.best <- regsubsets(price ~ size + lot + bath + bedrooms + agestandardized + garagesize + status + elem, data = train, nvmax = 13)
test.matrix <- model.matrix(price ~ size + lot + bath + bedrooms + agestandardized + garagesize + status + elem, data=test, nvmax = 13)


valid.errors <- rep(NA, 13)

i = 0
for(i in 1:13){
  coef.reg=coef(reg.best,id=i)
  pred=test.matrix[,names(coef.reg)]%*%coef.reg
  valid.errors[i] <- mean((test$price-pred)^2)
}

valid.errors
which.min(valid.errors) #Which model with the smallest error
```

```{r}
reg.best <- regsubsets(price ~ size + lot + bath + bedrooms + agestandardized + garagesize + status + elem, data = house)
coef(reg.best,which.min(valid.errors))
```

```{r}
model3 <- lm(price ~ size + lot + bedrooms + status + elem, data = house)
summary(model3)
```
  - This model has a good multiple r-squared value of 0.5277 and F-statistic, 7.263. But the adjusted r-squared value is smaller than the model from the previous part and from project 1. 
  
# E: Ridge Regression 
```{r}
train.x = model.matrix(price ~ ., train)
train.y = train$price
test.x = model.matrix(price ~ ., test)
test.y = test$price

set.seed(1)

cv.out = cv.glmnet(train.x, train.y, alpha = 0)
plot(cv.out)
```
```{r}
bestlam = cv.out$lambda.min
bestlam

RidgeModel <- glmnet(train.x, train.y, alpha = 0, lambda = bestlam, thresh = 1e-12)
coef(RidgeModel)

model4 <- RidgeModel
```

#F: Principal Components Regression/Partial Least Squares Regression
```{r}
set.seed(1)
pcr.fit=pcr(price ~ ., data=house,scale=TRUE,validation="CV")
pcr.min <- which.min(RMSEP(pcr.fit)$val[1,,]) - 1
validationplot(pcr.fit,val.type="MSEP")
```

```{r}
set.seed(1)
pls.fit=plsr(price ~ ., data=house,subset = index ,scale=TRUE, validation="CV")
summary(pls.fit)
validationplot(pls.fit,val.type="MSEP")
```

- After comparing both PLS and PCR we choose the PCR model. The PCR cross validation error is minimized at `r pcr.min` components as shown in the plot. This model had slightly better overall comparative values versus the PLS model.
  
#G: Errors and Comparisons

```{r}
Pred1 <- predict(model1, test)
Error1 <- mean((Pred1-test.y)^2)

Pred2 <- predict(model2, test)
Error2 <- mean((Pred2 - test.y)^2)

Pred3 <- predict(model3, test)
Error3 <- mean((Pred3 - test.y)^2)

Pred4 = predict(model4, s = bestlam, newx = test.x)
Error4 <- mean((Pred4-test.y)^2)

# We were stating PCR number of components, yet using the PLS model
# Pred5 = predict(pls.fit, test, ncomp = 1)
# Error5 <- mean((Pred5-test.y)^2) 

Pred5 = predict(pcr.fit, test, ncomp = 7)
Error5 <- mean((Pred5-test.y)^2)
 
cbind(Error1, Error2, Error3, Error4, Error5)
```

```{r}
rsq = function (predicted, actual) {
  return(1 - (sum((predicted - actual) ^ 2)/sum((actual - mean(actual)) ^ 2) ))
}

RSQ1 <- rsq(Pred1, test.y)

RSQ2  <- rsq(Pred2, test.y)

RSQ3  <- rsq(Pred3, test.y)

RSQ4 <- rsq(Pred4, test.y)

RSQ5  <- rsq(Pred5, test.y)
cbind(RSQ1, RSQ2, RSQ3, RSQ4, RSQ5)
```

  - Based on the errors, we can rank our canidate models as follow(smaller to larger error vlaues):
    1. Model 1 / Inital Model = 1832.57
    2. Model 2 / Regsub Model over all data = 2068.742
    3. Model 5 / PCR Regression = 2085.447
    4. Model 3 / Regsub Model over training data = 2197.164
    5. Model 4 / Ridge Regression = 2574.623
    
  - We can rank the models based on their $R^2$ values as follows: 
    1. Model 1 / Initial Model = 46.33
    2. Model 2 / Regsub Model over all data = 39.41
    3. Model 5 / PCR Regression = 38.92
    4. Model 3 / Regsub Model over training data = 35.65
    5. Model 4 / Ridge Regression = 24.60
    
# Final Model Presentation
  - We decided to stick with our inital model, the final model from Project 1 as our final model for Project 2 based on it having the lowest MSE and highest $R^2$ value computed above. 
  
```{r}
FinalModel <- lm(price ~ bedrooms + bath + garagesize + status + 
                       elem + size + lot + bedrooms:bath, house)
summary(FinalModel)
```

  - This final model accounts for aproximately 57% of the variation and the lowest MSE (1832.57). 
  
  - Model 2 is a good runner up model, the multiple r-sqaured value slightly descreases in comparison to model1 / Final Model, but the f-statistic slightly increases.
  
```{r}
summary(model2)
```

